---
title: "Generating BG & Category + SC Variables"
author: "Benet Ge"
date: "2023-07-10"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generating Block Groups From Lat/Lon

We will begin by reading in the data from the STATA file.

```{r libraries, include = FALSE}
library(sf)
library(dplyr)
library(haven)
library(readr)
library(data.table)
library(here)
```
```{r read data, cache=TRUE}
NYC311a2pared <- read_dta(here("NYC311datafiles", "pared data", "NYC311a2pared.dta"))
```

You can also then upload a shapefile for NYC block groups, downloaded from [here.](https://www2.census.gov/geo/tiger/GENZ2018/shp/)

```{r upload shapefile}
block_groups <- st_read(here("NYCshapefiles", "cb_2018_36_bg_500k", "cb_2018_36_bg_500k.shp"),
                        quiet = TRUE)
```
Then view a sample header of the data to make sure it's uploaded correctly.
```{r view initial data, cache=TRUE}
head(NYC311a2pared, n=10)
```

Next, filter out the calls for which no latitude and longitude were recorded. 

Draw coordinates from the latitude and longitude columns, and then convert them to shapefile objects. 

Retrieve the coordinate reference system from block groups.

Then run a function to assess the amount of time it would take to match these coordinates to block groups.
```{r generate the lat/lon conversion function}
latlong_sf <- NYC311a2pared %>%
  filter(!is.na(latitude), !is.na(longitude)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = st_crs(block_groups))

system.time({
  intersected <- st_intersects(latlong_sf, block_groups)
})

```

Create new columns that assign the block group that's matched to the coordinate pairs.
```{r mutate and assign block groups}
latlong_final <- latlong_sf %>% 
  mutate(intersection = as.integer(intersected),
         fips = if_else(is.na(intersection), "",
                        block_groups$GEOID[intersection]))
```
View the results and verify that this has been done correctly.
```{r verify the results}
latlong_relevant <- latlong_final[,c(1,46,47)]
head(latlong_relevant, n=10)
```


