---
title: "Analysing U.S. Census Data with R"
author: "Benet Ge"
date: "2023-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(tigris)
library(tidycensus)
library(tidyverse)
library(mapboxapi)
library(tmap)
```


# ggplot method

```{r cars}
hennepin_race <- get_decennial(
  geography = "tract",
  state = "MN",
  county = "Hennepin",
  variables = c(
    Hispanic = "P2_002N",
    White = "P2_005N",
    Black = "P2_006N",
    Native = "P2_007N",
    Asian = "P2_008N"
  ),
  summary_var = "P2_001N",
  year = 2020,
  geometry = TRUE
) %>%
  mutate(percent = 100 * (value / summary_value))
```
Now trying tmap method.
```{r}

hennepin_black <- filter(hennepin_race,
                         variable == "Black")
tm_shape(hennepin_black) +
  tm_polygons()
tm_shape(hennepin_black) +
  tm_polygons(col = "percent")
```



```{r}
tm_shape(hennepin_black) + # so we are now classing by "quantile"
  tm_polygons(col = "percent",
              style = "jenks",
              n = 5,
              palette = "Purples",
              title = "2020 US Census",
              legend.hist = TRUE) +
  tm_layout(title = "Percent Black\nby Census tract",
            frame = FALSE,
            legend.outside = TRUE,
            bg.color = "grey70",
            legend.hist.width = 5,
            fontfamily = "Verdana")
```



Adding reference elements to the map using mapboxapi






```{r static tiles}
hennepin_tiles <- get_static_tiles(
  location = hennepin_black,
  zoom = 10,
  style_id = "light-v9",
  username = "mapbox"
)
```




```{r}
tm_shape(hennepin_tiles) +
  tm_rgb() +
  tm_shape(hennepin_black) +
  tm_polygons(col = "percent",
              style = "jenks",
              n = 5,
              palette = "Purples",
              title = "2020 US Census",
              alpha = 0.7) +
  tm_layout(title = "Percent Black\nby Census tract",
            legend.outside = TRUE,
            fontfamily = "Verdana") +
  tm_scale_bar(position = c("left", "bottom")) +
  tm_compass(position = c("right", "top")) +
  tm_credits("(c) Mapbox, OSM    ",
             bg.color = "white",
             position = c("RIGHT", "BOTTOM"))
```



```{r}

tm_shape(hennepin_tiles) +
  tm_rgb() +
  tm_shape(hennepin_black) +
  tm_polygons(col = "percent",
              style = "jenks",
              n = 5,
              palette = "Purples",
              title = "2020 US Census",
              alpha = 0.7) +
  tm_layout(title = "Percent Black\nby Census tract",
            legend.outside = TRUE,
            fontfamily = "Verdana") +
  tm_scale_bar(position = c("left", "bottom")) +
  tm_compass(position = c("right", "top")) +
  tm_credits("(c) Mapbox, OSM    ",
             bg.color = "white",
             position = c("RIGHT", "BOTTOM"))
```

## 6.3.4: Alternative map types with tmap
Choropleth maps best suited for visualizing: 

* rates
* percentages
* statistical values that are normalized for the population of areal units

Not ideal for:

* comparing counts themselves

Because those maps may just reflect the underlying population's size.


### 6.3.4.1: Graduated symbols
`tm_bubbles()` can create a graduated symbol map of the Black population in Hennepin County:

```{r Graduated symbol map}
tm_shape(hennepin_black) +
  tm_polygons() +
  tm_bubbles(size = "value", alpha = 0.5,
col = "navy",
             title.size = "Non-Hispanic Black - 2020 US Census") +
  tm_layout(legend.outside = TRUE,
            legend.outside.position = "bottom")
```

This allows comparisons between circles & not polygons.

### 6.3.4.2: Faceted maps

For comparisons, `tm_facets()` creates a faceted map that identifies unique groups, with facets defined with `by`.


```{r Faceted map}
tm_shape(hennepin_race) +
  tm_facets(by = "variable", scale.factor = 4) +
  tm_fill(col = "percent",
          style = "quantile",
          n = 6,
          palette = "Blues",
          title = "Percent (2020 US Census)",) +
  tm_layout(bg.color = "grey",
            legend.position = c(-0.7, 0.15),
            panel.label.bg.color = "white")
```
Drawbacks:

* within-group variation/heterogeneity suppressed

### 6.3.4.3: Dot-density maps

`as_dot_density()` from `tidycensus` helps prepares Census data for dot-density. 

```{r as_dot_density}
hennepin_dots <- hennepin_race %>%
  as_dot_density(
    value = "value",
    values_per_dot = 100,
    group = "variable"
)
```
Paying attention above to the options.

`tm_dots()` creates the map.

```{r dot-density map}
background_tracts <- filter(hennepin_race, variable == "White")
tm_shape(background_tracts) +
  tm_polygons(col = "white",
              border.col = "grey") +
  tm_shape(hennepin_dots) +
  tm_dots(col = "variable",
          palette = "Set1",
size = 0.005,
          title = "1 dot = 100 people") +
  tm_layout(legend.outside = TRUE,
            title = "Race/ethnicity,\n2020 US Census")
```
As a note, a dasymetric dot-density map removes uninhabited polygons and `erase_water()` from `tigris` removes bodies of water.

## 6.4: Cartographic workflows with non-Census data
When using data that's avialable through Census geography but not through ACS/decennial (as in 311 calls), geometry = TRUE functionality not possible. 

### 6.4.1: National election mapping with tigris shapes

Interlude from [here](https://search.r-project.org/CRAN/refmans/tigris/html/block_groups.html)

```{r testing the interlude, results='hide'}
benton_bgs <- block_groups("Oregon", "Benton")



```
```{r plotting the geometry}
plot(benton_bgs$geometry)
```

Okay now practicing this with 311 and forlove (counts) data. *NOTE:* `left_join()` retains all rows from first dataset (likely the tigris, which is all bgs), whereas `inner_join()` drops non-matching rows in the output dataset. Try both every time.

### 6.4.2: Understanding and working with ZCTAs
```{r zcta data reading}
irs_data <- read_csv("https://www.irs.gov/pub/irs-soi/18zpallnoagi.csv")

ncol(irs_data)
```
`select()` allows you to show only certain variables.
```{r self-employment percentage}
self_employment <- irs_data %>%
  select(ZIPCODE, self_emp = N09400, total = N1)
```

```{r obtain ZCTAs for near Boston, MA}
library(mapview)
library(tigris)
options(tigris_use_cache = TRUE)
boston_zctas <- zctas(
  cb = TRUE,
  starts_with = c("021", "022", "024"),
  year = 2018
)
```

```{r view Boston ZCTAs}
mapview(boston_zctas)
```
```{r joining IRS dataset with ZCTA shapefiles}
boston_se_data <- boston_zctas %>%
  left_join(self_employment, by = c("GEOID10" = "ZIPCODE")) %>%
  mutate(pct_self_emp = 100 * (self_emp / total)) %>%
  select(GEOID10, self_emp, pct_self_emp)
```

```{r actually generating ZCTA IRS map}
tm_shape(boston_se_data, projection = 26918) +
  tm_fill(col = "pct_self_emp",
          palette = "Purples",
          title = "% self-employed,\n2018 IRS SOI data")
```
But if instead we want to visualize where most self-employment is located, instead of a percentage relative to the area, we'd want to use 'self_emp' itself in a bubble diagram:
```{r IRS bubble ZCTA data}
tm_shape(boston_se_data) +
  tm_polygons() +
  tm_bubbles(size = "self_emp",
             alpha = 0.5,
             col = "navy",
             title.size = "Self-employed filers,\n2018 IRS SOI data")
```

## 6.5 Interactive mapping

### 6.5.1: Interactive mapping with Leaflet
`leaflet` package does interactive mapping 



```{r setting up a leaflet color palette}
library(leaflet)
pal <- leaflet::colorNumeric(
  palette ="magma",
  domain = nyc_bgs_joined_all_wgs$"0"
)

```


```{r building the leaflet map}
leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = nyc_bgs_joined_all_wgs,
              color = pal(nyc_bgs_joined_all_wgs$"0"),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.5,
              label = nyc_bgs_joined_all_wgs$"0") %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = nyc_bgs_joined_all_wgs$"0",
    title = "# type 0 complaints NYC all"
  )
```
But we lose in group variation because of outliers. We do this with the function `colorQuantile()` instead of `colorNumeric`. Create a new pal called `pal_q`.

```{r setting up a quantile color palette, include = FALSE}
library(leaflet)
pal_q <- leaflet::colorQuantile(
  palette ="magma",
  domain = nyc_bgs_joined_all_wgs$"0"
)
```

```{r building the leaflet map 2}
leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = nyc_bgs_joined_all_wgs,
              color = pal_q(nyc_bgs_joined_all_wgs$"0"),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.5,
              label = nyc_bgs_joined_all_wgs$"0") %>%
  addLegend(
    position = "bottomright",
    pal = pal_q,
    values = nyc_bgs_joined_all_wgs$"0",
    title = "# type 0 complaints NYC all"
  )
```

```{r troubleshooting the crs inequivalence, eval = FALSE}
nyc_bgs_joined_all_wgs <- st_transform(nyc_bgs_joined_all, crs = "WGS84")
```

This is totally awesome. The thing above is to document a difficulty because the CRS used by the tidycensus c_bgs was NAD83, and `leaflet` wants WGS84. Some things to still adjust, however:

* Notice how it's difficult to identify boundaries between contiguous block groups within the same percentile.
* In addition to the legend including our percentiles, we want to display what those actual percentiles are.

```{r troubleshooting the boundary issue}
leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = nyc_bgs_joined_all_wgs,
              color = pal_q(nyc_bgs_joined_all_wgs$"0"),
              weight = 0.5,
              smoothFactor = 1.0,
              fillOpacity = 0.5,
              label = nyc_bgs_joined_all_wgs$"0") %>%
  addLegend(
    position = "bottomright",
    pal = pal_q,
    values = nyc_bgs_joined_all_wgs$"0",
    title = "# type 0 complaints NYC all"
  )
```
That's solved. But what about the quantiles? Find the actual values here:
```{r quantiles}
nyc_bgs_2012_quantiles <- quantile(nyc_bgs_joined_all_wgs$"0", na.rm = TRUE)
print(nyc_bgs_2012_quantiles)
```
Then apply that to the leaflet code:
```{r}
leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = nyc_bgs_joined_all_wgs,
              fillColor = pal_q(nyc_bgs_joined_all_wgs$"0"),
              color = "black",
              stroke = TRUE,
              weight = 0.3,
              smoothFactor = 0,
              fillOpacity = 0.9,
              label = nyc_bgs_joined_all_wgs$"0") %>%
  addLegend(
    position = "bottomright",
    colors = c('black', 'purple', 'red', 'yellow'),
    values = nyc_bgs_joined_all_wgs$"0",
    labels = c("Q1: 0-10", "Q2: 11-26", "Q3: 27-53", "Q4: 54-1559"),
    title = "# 2012-2014 type 0 complaints NYC"
  )
```

Somehow I tried bins with Jenks generated numbers:

```{r setting up Jenks bins}
install.packages("BAMMtools")
nyc_bgs_2012_jenks <- BAMMtools::getJenksBreaks(nyc_bgs_joined_all_wgs$"0", 5)
print(nyc_bgs_2012_jenks)

bins <- c(0,56,164,439,1559)
pal_bins <- colorBin("YlOrRd", domain = nyc_bgs_joined_all_wgs$"0", bins = bins)
```

```{r Jenks bins}
leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = nyc_bgs_joined_all_wgs,
              fillColor = pal_bins(nyc_bgs_joined_all_wgs$"0"),
              color = "black",
              stroke = TRUE,
              weight = 0.3,
              smoothFactor = 0,
              fillOpacity = 0.9,
              label = nyc_bgs_joined_all_wgs$"0") %>%
  addLegend(
    position = "bottomright",
    colors = c('#FFFFB2', '#FECC5C', '#FD8D3C', '#E31A1C'),
    values = nyc_bgs_joined_all_wgs$"0",
    labels = c("J1: 0-55", "J2: 56-163", "J3: 164-438", "J4: 439-1559"),
    title = "# 2012-2014 type 0 complaints NYC"
  )
```

```{r Jenks with interactivity}
leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = nyc_bgs_joined_all_wgs,
              fillColor = pal_bins(nyc_bgs_joined_all_wgs$"0"),
              color = "black",
              stroke = TRUE,
              weight = 0.3,
              smoothFactor = 0,
              fillOpacity = 0.9,
              highlightOptions = highlightOptions(
                weight = 3,
                color = "#F8C",
                fillOpacity = 0.7,
                bringToFront = TRUE),
              label = labels_jenks,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
  addLegend(
    position = "bottomright",
    colors = c('#FFFFB2', '#FECC5C', '#FD8D3C', '#E31A1C'),
    values = nyc_bgs_joined_all_wgs$"0",
    labels = c("J1: 0-55", "J2: 56-163", "J3: 164-438", "J4: 439-1559"),
    title = "# 2012-2014 type 0 complaints NYC"
  )
```

